### 2014-03-15 paper.tex
revised according to conversation with Eric Barton

\documentclass[conference]{sig-alt-gov2}
\pdfpagewidth=8.5in
\pdfpageheight=11in
%\usepackage{subfig}
\usepackage{subfigure}
\usepackage[pdftex]{}
\usepackage{todonotes}
\usepackage{listings}
\lstset {% general command to set parameter(s)
         language=C,
     basicstyle=\footnotesize,               % print whole listing small
%     keywordstyle=\color{black}\bfseries, % underlined bold black keywords
%     identifierstyle =\color{black},  % nothing happens
%     commentstyle=\color{black}\emph, % white comments
     %stringstyle=\ttfamily,          % typewriter type for strings
%     stringstyle=\color{black},       % typewriter type for strings
         tabsize=4,
         showtabs=false,
     showstringspaces=false}
%can't figure this one out for particles bandwidth
%\usepackage[caption,label]{subfig}

%\newcommand{\TITLE}{D\textsuperscript{\huge{2}}T: Doubly Distributed Transactions for High Performance and Distributed Computing}

\newcommand{\DDT}{D\textsuperscript{2}T~}
\newcommand{\DDTns}{D\textsuperscript{2}T}

\hyphenation{sub-trans-ac-tion}

% in order for balance columns to work, it has to be before begin document...
%\balancecolumns

\begin{document}

\conferenceinfo{MSST'14,} {November 17, 2013, Denver, Colorado, USA.}
\CopyrightYear{2014}
%\crdata{978-1-4503-1103-8/11/11}
\clubpenalty=10000
\widowpenalty = 10000

\title{Consistency and Fault Tolerance Considerations for the Next Iteration of the DOE Fast Forward Storage and IO Project}

\numberofauthors{3}
\author{
\alignauthor Jay Lofstead\\
       \affaddr{Sandia National Laboratories}\\
       \email{gflofst@sandia.gov}
\alignauthor Ivo Jimenez\\
       \affaddr{University of California, Santa Cruz}\\
       \email{ivo@cs.ucsc.edu}
\alignauthor Carlos Maltzahn\\
       \affaddr{University of California, Santa Cruz}\\
       \email{carlosm@soe.ucsc.edu}
%\and  % use '\and' if you need 'another row' of author names
}
\maketitle

\begin{abstract}
With phase 1 of the Fast Forward Storage and IO Stack project complete, it is
an excellent opportunity to evaluate many of the decisions made to feed into
the phase 2 effort. The initial effort to define a next generation file system
has made admirable contributions in architecture and design. Formalizing the
general idea of data staging as burst buffers for the file system will help
manage the performance variability and offer the data processing opportunities
outside the main compute and file system. Adding a transactional mechanism to
manage faults and data visibility helps enable effective analytics without
having to work around the IO stack semantics.

While these and other their contributions are valuable, similar efforts made
elsewhere may offer attractive alternatives or differing semantics that would
yield a more feature rich environment with little to no additional overhead.
For example, the Doubly Distributed Transactions (\DDTns) protocol offers an
alternative approach for incorporating transactional semantics into the data
path. The PreDatA project examined how to get the best throughput for data
operators can offer some insights for refining the Burst Buffer concept.

This paper examines some of the choices made by the FastForward team and
compares them with other options and offers observations and suggestions based
on these other efforts.  This will include some non-core contributions of other
projects, such as some of the demonstration metadata and data storage
components generated while implementing \DDTns, to make suggestions that may
help the next generation design for how the IO stack works as a whole.

\end{abstract}

\category{D.4}{Software}{Operating Systems}
\category{D.4.7}{Operating Systems}{Organization and Design}[hierarchical design]

\terms{Design, Performance}

\section{Introduction}

Current production HPC IO stack design is unlikely to offer sufficient features
and performance to adequately serve the needs of an extreme scale platform. To
address these limitations, the US Department of Energy commissioned an effort
to develop a design and prototype for an IO stack suitable for the extreme
scale environment. This is a joint effort led by the Intel Lustre team, Los
Alamos National Laboratory, EMC, DDN, and the HDF Group. This team has
developed a specification set~\cite{fastforward:2014:docs} for a future IO
stack to address the identified challenges. The first phase recently completed
with a second phase getting underway. The core focus of the first phase was
basic functionality and design. The second phase will refine this design
incorporating fault recovery and other features missing from the first phase.

The overall design seeks to offer a byte-granular, multi-version concurrency
control. Through the use of a copy-on-write style mechanism, multiple versions
of a file can be stored in potentially greatly reduced space. By assuming the
client interface will be through an IO library affords a more complicated
interface that offers richer functionality without requiring user code changes.
By managing data in the IOD layer rather than requring writing to centralized
storage will better support the performnce requirements of integrated
application workflows. At a more detailed view, the various layers of the IO
stack each contribute different functionality and performance implications.

The basic architecture incorporates five layers. The top layer is a high level
IO library, such as the demonstration HDF-5 library. The intent is to only have
access to the storage stack through such an API to manage the complexity of
working with the lower layers. Below the user API is an IO forwarding layer
that redirects IO calls from the compute nodes to the IO dispatching layer.
This IO forwarding layer is analogous to the function of the IO nodes in a
BlueGene machine. The next two layers have considerable functionality. The IO
dispatcher (IOD) serves as the primary storage interface for the IO stack and
is the only way to reach the persistent storage array in lower layers.

The core idea for IOD is to provide a way to manage the IO load that is
separate from the compute nodes and the storage array. Communication intensive
activities, such as two-phase, data sieving IO's data rearrangement process,
can be moved to the IOD layer offloading the communication load from the
compute nodes. IOD has three main purposes. First, if the optional burst buffer
is available, it works as a fast cache absorbing write operations for the
slower trickle out to the central storage array. It can also be used to
retrieve objects from the central storage array for more efficient read
operations and offer data filtering to make client reads more efficient.
Second, it offers the transaction mechanism for controlling data set visibility
and to manage faults that would prevent a data set from being used. Third, data
processing operations can be placed in the IOD. These operations are intended
to offer data rearrangement, filtering, and similar operations prior to it
reaching the central storage array.

While these ideas are not necessarily new, they are new twists on best of class
efforts for these techologies. For example, offloading the collective two-phase
data sieving from the compute nodes to reorganize data has proven effective at
reducing the total time for writing data due to fewer participants involved in
the communication patterns~\cite{lofstead:2011:nessie-staging}.  Beyond these
broad items, there are many important details some of which are examined in
more detail below.

The bottom two layers are the Data Access Object Storage (DAOS) and Versioning
Object Storage Device (VOSD). DAOS is the typical interface that represents
the individual objects stored including the semantics related to ``files''.
VSOD is the actual storage mechanism for the objects defined at the DAOS layer.

The focus of this paper is primarily the IOD layer given the critical role it
has in the performance and functionality of the entire stack. Most of the key
features explored in this paper all have a strong presence in the IOD layer
motivating the focus of this examination.

Along with the analysis of the published design documents, a discussion of the
design philosophy representing the overall intent is presented. This informtion
reprents information that may or moy not have been written down, but is the 
intent of ultimate product. These insights were gained based on personal
converations with the team members discussing some of the potential challenges
with the design as written. These ideas are presented to give a fuller picture
of where the project is going rather than dwelling on the limitations of the
published documents.

The rest of the paper is organized as follows. Section~\ref{sec:burst}
discusses some of the features of incorporating burst buffers as designed and
suggests some considerations and alternatives for the next generation of this
project. Section~\ref{sec:transactions} discusses the transactions approach
offered in the IOD layer and the corresponding epochs in the DAOS layer. It
also offers a comparison to the \DDT system given the very similar high-level
design and motivating use case.  Section~\ref{sec:summary} discusses the system
overall with recommendations on what design elements should be considered based
on broader issues with current HPC data centers. The paper is concluded in
Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
paper.

\section{Burst Buffers}
\label{sec:burst}

The idea of burst buffers were initially explored in the context of data
staging~\cite{abbasi:2007:datatap,Abbasi:2009:datatap,nisar:2008:staging,zheng:2010:predata}.
These initial designs all use extra compute nodes to represent the data storage
buffer given the lack of any dedicated hardware support for this functionality.
The desired outcome of these initial studies is to motivate how such
functionality might be incorporated and the potential benefits.  Later, these
concepts were proposed to be incorporated as part of the IO
stack~\cite{bent:2012:challenges,bent:2012:burst-buffer}.  The current
FastForward IOD design recommends incorporating SSDs, but specifically lists
these devices as optional. Unfortunately, both incorporating burst buffers and
the use of SSDs in the IOD layer may be problematic.  First, with these burst
buffers being optional, the semantics must be clarified of how IOD works must
change to use DAOS to store the data directly rather than storing them locally
until explicitly persisted by the user. The design of DAOS does not incorporate
a high transaction rate mechanism. Conversely, it assumes that it will only be
involved when persisting a completed transaction and only for a fraction of the
total transactions created. Further, with function shipping offering the
ability to change how the data is stored and arranged prior to it being written
to DAOS, this functionality may also be optional depending on the existence of
the burst buffer or slow the IO path to accommodate the data processing.
Consider the important functionality of data rearrangement and doing things
like changing the fast array dimension.

One of the bigger concerns is the observation that the original data staging
proposals all used compute nodes while the newer proposals seek not only to
make them a fixed portion of the IO stack, but also shared across all machine
users. The PreDatA~\cite{zheng:2010:predata} paper in particular examines the
potential costs and advantages of where to place operators similar to the IOD
proposed function shipping. There are two key takeaways from that work. First,
placement matters.  Depending on the communication intensity vs. computation
intensity, where along the IO path to place the operation can matter
significantly. Second, and more importantly, the amount of time spent
processing for the operators was stretched to the point where it consumed
nearly all of the time between IO operations. The given ratios of compute
processes to staging process examined is representative for future extreme
scale platforms. If anything, the ratios offer more staging processes than IOD
processes would be available.

In the case of the written IOD design, it forces a staging area deployment and
then shares the staging area across all users simultaneously. This is unlikely
to be useful because of the limited compute and communication capacity to spare
to perform these operations at a bottleneck in the IO path. The use of a
separate data staging area intentionally separate from the IO path allows using
operators on limited resources leaving the IO path clear for strictly data
movement.

By concentrating this functionality into the storage stack, three problems
arise.  First, the amount of network bandwidth, IO bandwidth, and compute power
consumed for example operations from a single application is likely to
completely monopolize the IOD processes. Second, if space and time partitioning
is used instead, the functionality risks being too small to be useful. Third,
the hardware performance advantage for SSDs is questionable. Current NAND-based
flash devices top out at around 400 MB/sec. The key spec that is missing from
this number is that 400 MB/sec is a measure of the fixed number of available
IOPS multiplied by the block size. This represents the ideal streaming
performance possible. The problem is that it costs an IOP to read 1 byte or 1
block (4KB or 8KB, depending on the device). It costs 1 IOP to write a full
block--usually. In some cases, it will cost 2 IOPs.  This accounts for the
required pre-erase write prior to writing to a reused block. In the worst case,
it can be 3 IOPS per write. This would be for a partial block write (read the
old block, erase, write the modified block). One-third of 400 MB/sec, about 133
MB/sec, is well below the streaming performance of HDDs.  Granted, there is
still rotational and seek latency to deal with for HDDs, but the advantage for
SSDs has evaporated and potentially turned into a penalty at a considerable
cost premium.  There are faster SSD solutions on the market that incorporate
DRAM for caching and using the PCIe bus, for example, but their price precludes
them from use in an extreme scale platform.

Given these features, the optionality and even incorporation of burst buffers
in the current design should be carefully considered. Much of the advanced, key
functionality proposed as they are currently designed ultimately relies on the
existence of burst buffers to work. Further thought about how to have an IOD
layer both with and without a burst buffer is required before they can be
considered optional. As the design stands today, they are a required part of
the IOD layer for proper functioning. Unfortunately, it is not clear that they
can address the performance concerns they are intended to cover.

\subsection{Design Philosophy}
The burst buffers design, as presented in the IOD documents, limits the
placement of the function operators and SSDs for buffers to the IO nodes. The
team does acknowledge the limitations of this design and intend to ultimately
focus on spreading the IOD from the IO nodes into the compute area as well.
This is intended both to help address the limitations of the IO bandwidth
and compute capaibility of these few nodes for data processing, but also to
take advantage of new layers in the storage hierarchy. By incorporating NVRAM
into compute nodes, new options for buffering data prior to being moved to
centralized storage become available. This lessens the impact of some operators
while offering additional options for places to store data.

Burst buffers being optional is a high level goal, but not one considered at a
detailed level with the design. While this forces some serious thought on how
to adapt the functionality proposed for IOD to work directly on DAOS will have
to be considered. With the additional desire to support using compute node
resources for these operations, serious work will be required to make a fully
functional end-to-end IOD layer implementation for a production system.

Another concern that is acknowledged, but no thought as been applied to is the
requirement that a single IOD process be the master for any operation. Should
the number of concurrent applications exceed the available nodes, sharing an
IOD process will be required. Since partitioning of the IOD processes for
exclusive use by particular applications is the proposed operating mode, should
insufficient IOD resources be available, either a job could be delayed or IOD
resources could be reallocated from a different process could be redeployed for
use by the new job. These sorts of consideraitons still need to be made for a
full production system.

\section{Transactions and Epochs}
\label{sec:transactions}

The transaction mechanism manifests in two forms. At the IOD layer, they are
called transactions and are used to judge whether or not an output is complete
or not and control access through treating the transaction ID as a version
identifier. At the DAOS layer, they are called epochs and represent persisted
(durable) transactions from the IOD layer. Each of these offers different
functionality, but are connected as is explained below. How these differ from
the \DDT approach is also explored. While IOD's and \DDTns's transactions are
seemingly very different, they use a similar high-level design, but very
different implementation, to solve the same problem.

\subsection{IOD Transactions}
To understand how transactions are used in the IOD layer, some terminology and
concepts must be explained first. At the coarsest grain level is a container.
Each container serves to host a transaction and contains a collection of
objects. Conceptually, containers corresponds to a file in a traditional file
system. The objects in each container represent different data within a file.
The three initially defined object types are key-value store, array shard,
and blob. The easiest way to understand these types is to evaluate these from
the perspective of an HDF-5 file, the initial user interface layer. The
key-value store represents a collection of attributes. The array shard
represents part of a potentially multi-dimensional array. It is referred to as
a shard because it is likely a small piece of a globally defined array. The
blob represents a byte stream of arbitrary contents.  The fundamental
difference between an array shard and a blob is that the array shard has
metadata identifying its portion within the global, logical space while the
blob is simply a 1-dimensional array of bytes that is not shared across IOD
nodes.  Should an operation be deployed into IOD to manipulate an array within
a container, it would operate on the array shards rather than on blobs. Given
this context, the transactions come in two forms.

First is a single leader transaction where the IOD manages based on calls from
a single client. The underlying assumption is that the client side will manage
the transactional operations itself and the single client is capable of
reporting to the IOD how to evolve the transaction state.

The second form is called multi-leader and has the IOD layer manage the
transactions. In this case, when the transaction is created, a count of clients
is provided to the IOD layer. As clients write to the container, the reference
count is reduced. Once the count reaches 0, the transaction is automatically
committed.

\subsubsection{Design Philosophy}
Undocumented, but inherent in the design of these transactions is how faults
are detected. The initial design assumes that the current Lustre fault
detection mechanism that can determine if a process or node is no longer
reachable is used. This defines how a fault will be detected and what will trigger a passive fault recovery (i.e., transaction abort).

A new term, ``Primary'', is used to define the process(es) that contact the IOD
layer directly to open a container. Unfortunately, this term is never used nor
defined within any of the published documents. There are some very positive and
negative performance implications of this approach. On the positive side, like
PVFS, it is possible for a single process to open a container and share that
handle with other processes reducing the load on the IOD layer. The key idea is
that the single leader vs. multi-leader transaction mode is automatically defined based on the number of Primary processes that open a container. This greatly
simplifies any additional client-side load in using the different styles of transactions. Also on the positive side, it offers the ability to have a subset of
processes that are considered transaction participants affording reduced load
on the IOD processes and more localized transaction management on the client
side. On the negative side, for a fully IOD managed transaction, ALL processes
must open the container forcing a potentially immense number of simultaneous
calls to the IOD layer to open a container. This sort of onslaught is what is
dealt with on metadata servers today with mass file opens. The reality is that
this ends up having the same sort of performance implications as distributed
denial of service attacks on the IOD processes.

Ultimately, with the passive detection of faults for Primaries, the transaction
mechanism can work very well. A mostly unstated restriction that will be
relaxed in the next phase is that every sequential transaction on a container
is considered dependent on the earlier transaction. Should one output be
delayed and the subsequent five succeed, when the delayed process finally
fails, all six transactions are rolled back. The thought of using this
mechanism to store subsequent checkpoint outputs in the same container to both
save space, but not care if one fails, cannot work in the current form. This
has been acknowledged and is planned on being relaxed in phase two.

\subsection{DAOS Epochs}
The Epoch mechanism differs from transactions. Instead of focusing on when a
particular output is complete, an epoch represents incremental persisted copies
of a container. To simplify the mapping between an IOD transaction and the DAOS
epochs, when an IOD transaction is persisted to DAOS, the IOD transaction ID is
the used as the epoch ID. The key difference is that at the DAOS layer, some
transaction (epoch) IDs will not be represented.

\subsection{Metadata Management}
Eliminating metaata management as a special case and instead treating it just
as data is a central design goal of the Fast Forward project. Unfortunately,
it is not completely achieved.

Eliminating metadata as a core component of a file system is not new. It has
been explored as part of the Light Weight File Systems
project~\cite{oldfield:lwfs}. In LWFS, the metadata service is explicitly
limited to a user task with the storage layer limited to data
storage/retrieval, authorization, and authentication. This approach proved
workable. Using this hybrid approach is less common and introduces other
issues.

IOD and DAOS both share a philosophy that they will have to maintain the
metadata about how objects are striped and where they are placed. It was
decided that the additional data to track the namespace as well was a small
enough additional increment that it was worth maintaining within the system.

\subsubsection{Design Philosophy}
While the metadata design is not fully defined, there are a few things that
are intended as part of this. For example, there is a standard, well known
container that is the system metadata. This includes the list of all other
containers. This container is treated like any other data in the system and
striped as appropriate at both the IOD and the DAOS layers. Unfortunately, this
still couples the metadata to a single object that must serialize access. If
the metadata, including information about striping and other data layout
operations were separated completely from the data path, more scalable
throughput could be achieved. The real challenge of this is introduced by the
the IOD, DAOS, and VOSD layers. Each of these requires some different metadata
storage and the migration is transparent to the user. Supporting fully
independent metadata with this model is difficult, if not impossible. Serious
thought on how to do this outside the data path effectively should be
considered for phase two.

The \DDT project created a simplistic model for a metadata service that has
different features. It is described below and has been published
previously~\cite{lofstead:2012:txn-metadata}. Some of the major differences
include the following. First, \DDTns's metadata service has a way to get a list
of the objects in the metadata service. Since this was intended as a test
system to examine \DDT performance, we did not put the effort into a making it
a robust system. While it does not address having a file equivalence like the
container concept, it does introduce a way to figure out what is available.
Stored along with this object description is a list of the global dimensions
for array objects. One observation made as part of creating this service is that
this is insufficient for many engineering codes that use non-regular meshes.
Second, for each object, a list of the equivalent to shards are kept with a
link back to the master object and the offset and size for each dimension for
that shard. This approach is not scalable because of the explosion in the
process count. Using a striping approach like IOD does is superior. Third,
there is no mapping to another layer with different object layout and counts.

Based on the lessons from the \DDT metadata service construction and the prior
experiments with LWFS, having a completely separate metadata service is
workable. Rather than making it a bottleneck in the IO path, it is another
service that users must interact with if they need those services. Otherwise,
it is 100\% out of the way. Users can manage everything by maintaining the
metadata including the list of objects themselves. However, there are drawbacks
to this 100\% client-side approach.

With a client-side only approach, there is a serious risk of the metadata
service and the object store getting out of sync.  While having a metadata-less
object storage service is desirable, the different semantics from traditional
file systems requires some considerations. In this case, should these services
get out of sync, three particular risks introduced.  First, a client could
create an entry in the metadata service that does not correspond to any objects
in the object store. Second, a client could create objects that have no
associated metadata entries. Third, updates to the metadata or object store
service should be an atomic operation, but due to a lack of coordination, a
window where the system is inconsistent appears.  Ultimately, the consistency
semantics required must be determined. If a metadata service is required and it
must be in sync with the object storage service, then additional work must be
performed. In traditional file systems, the metadata and object storage updates
are atomic. With decoupling metadata from object storage, should this atomicity
still be desired, it requires both the ability for the services to participate
in a task that is part of a larger atomic operation and a higher-level
mechanism to manage the atomic operation. Overall, while additional work is
required to maintain a client-side only metadata service, eliminating any
potential bottlenecks related to updating metadata related to the object
storage are eliminated. The burden of tracking striping and other metadata that
has traditionally been part of the metadata associated with the file system
will have to be maintained by the object storage service. The lack of a
centralized, serialized bottleneck to store that information improves
concurrency.

\subsection{Comparison to \DDTns}
The \DDT project~\cite{lofstead:2012:txn} sought to develop an efficient
approach for handling ACID-style transactions in an environment with parallel
clients and multiple servers (doubly distributed). Rather than being aimed
solely at data movement operations, \DDT seeks to address the general problem
of managing any operation with multiple clients and servers.  Consider the
management of the analysis/visualization area, potentially similar to the IOD
concept. The transaction protocol is used to help manage resizing of the
resource allocation to the various analysis and visualization components.  For
the purposes of this discussion, \DDT could also be used to manage changing how
IOD processes and/or nodes are used without exposing these changes to the
client processes prematurely.  This has been described and analyzed
previously~\cite{dayal:2013:io-containers}.

The example metadata and data storage services created as part of the \DDT
project lack one of the key transaction management features used by IOD.
The major difference is that there are no dependencies between transactions
that prevent visibility should an older version be incomplete. This additional,
intentional requirement by IOD offers different functionality than \DDTns's
example services. In the case of \DDTns, the functionality is more minimal, but
also avoids some of the concerns outlined below.

The second iteration of the protocol~\cite{lofstead:2013:pdsw-txn} fixed
scalability issues and demonstrated a scalable client-side coordination model
with excellent performance. The performance measured for a complex transaction
with \DDT is illustrated in Figure~\ref{fig:performance}. This performance is
explored in detail in a previous paper~\cite{lofstead:2013:pdsw-txn}. The
breakdown of the number of participants in each role is shown in
Table~\ref{tab:scaling}. For comparison, consider the Number of
Sub-Coordinators equivalent to an IOD process. The Processes Per
Sub-Coordinator represents the number of clients that use a particular IOD. For
these tests, we maintained a balanced distribution and always used at least two
sub-coordinators to slow down the processing.

\begin{figure}[ht]
%\vspace{-0.15in}
\centering
\includegraphics[keepaspectratio=true, width=0.9\columnwidth]{images/performance}
\vspace{-0.15in}
\caption{Total Transaction Overhead}
\label{fig:performance}
\vspace{-0.15in}
\end{figure}

\begin{table}[ht]
    \vspace{-0.15in}
    \centering
    \caption[Scaling Configuration]{Performance Tests Scaling Configuration}
    \bigskip
    \vspace{-0.15in}

    \begin{tabular}{|r|r|r|}
\hline
Processes & \vtop{\hbox{\strut Number of}\hbox{\strut Sub-Coordinators}} & \vtop{\hbox{\strut Processes Per} \hbox{\strut Sub-Coordinator}}\\
\hline
%8 & 2 & 4 \\
%16 & 2 & 8 \\
%32 & 2 & 16 \\
%64 & 2 & 32 \\
%128 & 2 & 64 \\
256 & 2 & 128 \\
512 & 2 & 256 \\
1024 & 4 & 256 \\
2048 & 8 & 256 \\
4096 & 16 & 256 \\
8192 & 32 & 256 \\
16384 & 64 & 256 \\
32768 & 128 & 256 \\
65536 & 256 & 256 \\
\hline
    \end{tabular}
    \label{tab:scaling}
\end{table}

At a high level, both \DDT and the IOD transactions have the same high-level
design. In both cases, a hierarchical model is employed. In the case of \DDTns,
it is a purely client-side tree using semi-synchronous messaging. The messaging
itself, in the current implementation, uses asynchronous MPI messages. The
synchronous component comes from the timeout mechanism used to detect faults.
It forces a level of coordination and synchronization for the protocol. For
IOD, it is a server-side tree and fully asynchronous. In both cases, there is a
master in charge of managing the transaction and a collection of workers that
aggregate into the master through second-level leaders. Beyond that, there are
some significant differences. Some of the different choices made by IOD raise
some possible concern.

First, should a transaction on a container not complete, all subsequent
transactions on that container, even if they are complete and correct, will not
be accessible. Once the transaction fails, all of these subsequent transactions
will be aborted. While this is being reconsidered for phase two, it does
complicate the interface to offer the additional transaction mode.

Second, in the multi-leader model, using a count of client connections to
determine if transaction is complete is problematic.  First, should this be in
an environment where lost or repeated network messages are possible, then the
count could be wrong inappropriately. This is a general problem and unlikely
in the current environment, but is a concern on less integrated platforms.
The idea of a count precludes any tracking of expected messages from sources
introducing a potential hole. Second, the aggregation of completion messages
are based on the local aggregation point reporting to the transaction
leader the state.  There is no ability to detect or recover from a failure of
this IOD process that would report to the transaction leader. Third, because a
simple count of participating processes is sent with the beginning of the
transaction, each process is strictly limited to a single output operation.
This precludes a process writing multiple array shards, for example. To put
this in concrete terms, it means that a file could only contain a single,
globally distributed variable (at the root level of HDF5) OR an attribute per
process. If this count actually represents a begin/end transaction pair
instead, it suffers from a lack of accountability for missing object (shard)
writes. All it would indicate is that a process started and ended a transaction
connection. There is no validation that a process has done everything required
of it.

\begin{figure}[ht]
\centering
\includegraphics[keepaspectratio=true, width=0.9\columnwidth]{images/optimized-protocol}
\vspace{-0.15in}
\caption{Optimized Protocol}
\label{fig:optimized-protocol}
\vspace{-0.25in}
\end{figure}

\DDT has addressed this count issue in a couple of ways. First, the
sub-coordinators each have a list of processes from which they expect messages.
Should a message be missed, it is noticed and corrective action can be taken.
Second, \DDT has the concept of sub-transactions. The messaging requirements
are illustrated in Figure~\ref{fig:optimized-protocol}.  Sub-transactions
represent finer grained operations than the entire output, \DDT can manage
multiple writes per client by using a sub-transaction to represent the output
for any item to the file (container). Because of how the sub-transactions are
managed, the singleton sub-transactions, ones in which only a single process
participates, must be declared before the transaction begins so that its
existence can be broadcast as part of the begin transaction message. This
ensures there is global knowledge that the sub-transaction is expected.  That
way if the coordinator (transaction leader) fails, which ever process takes
over that role knows to expect a completion message for that sub-transaction or
the overall transaction cannot complete. Global sub-transactions can be defined
at any time since they are a global, synchronized operation broadcasting their
existence. While this additional layer does introduce messaging, the overhead
is quite small.

The advantages of eliminating these messages is not performance as demonstrated
by the performance of \DDTns. Instead, it offers a much less synchronous model
that matches with different programming models, such as Charm++ or other
task-based approaches. Since it can work for the bulk-synchronous model also,
it is a more broadly applicable approach. This assumes that the observed
potential issues can be addressed successfully.

\subsection{Comparison to Other Protocols}
Alternatives, such as Paxos~\cite{Lamport:1998:paxos} algorithms like
ZooKeeper~\cite{Hunt:2010:zookeeper}, suffer from two limitations making them
unsuitable for this environment. First, the distributed servers in Paxos
systems are all distributed copies of each other that eventually become
consistent. Given the scale we wish to address, a single node's memory is
unlikely to be able to hold all of the data necessary for many operations at
scale. They also do not have a guarantee for when consensus will be achieved
without using slower synchronous calls. For the tight timing we wish to
support, we need guarantees of when a consistent state has been achieved.
Second, these systems also all assume that updates are initiated from a single
client process rather than a parallel set of processes as is the standard in
HPC environments.

\DDT uses a second layer of coordination on the client side that greatly
increases the scalability by consolidating messages from clients into unique
sets prior to sending to the overall coordinator. A gossip
protocol~\cite{ganesh:2003:gossip-protocols} may appear sufficient for this
purpose, but the delay of eventual consistency is strictly avoided with this
protocol to ensure guarantees at particular states in the code. For example, if
a failure occurs, the global understanding of the role of all processes is
required in order for effective communication to occur for operations like
creating sub-transactions or voting. In this case, the protocol can offer
stronger statements about consistency than these protocols offer.  These
features offer a way to easily scale the transaction protocol given the
guarantees we wish to offer. The IOD approach does nothing to address these
concerns.

Another effort to offer consistency and data integrity for the ZFS file
system~\cite{zhang:2010:zfs} covers some of the same territory. Instead of a
focus on the processes all having a notion of completion as a transaction, this
work focuses on the integrity of the data movement operations. We view this
work as something that should be considered hand-in-hand with a transaction
approach to ensure the integrity of the movement of the data in addition to the
agreement of processes about the successful completion of a parallel operation.

\section{Broader Design}
\label{sec:summary}

At a broader level, there are some concerns that were partially clarified
through conversations with the team.  Consider a shared file system across an
HPC data center. The current design maintains the metadata in its own
container. Since copying data from the IOD layer to the DAOS layer requires an
explicit persist call, how and when synchronizing the metadata across the 
layers and potentially across machines occurs is undefined. By delaying the
synchronization until an explicit persist is called will reduce the update
frequency, it does delay the visibility of data. Ideally, the metadata object
would need to be automatically persisted every time a container transaction is
persisted to the DAOS layer.

The implication of this is that every transaction persist is double operation
to account for the metadata persist. More importantly, the IOD-layer version of
the metadata container may contain readable transactions that have not been
persisted to the DAOS layer. How to handle this inconsistency between the two
layers still needs to be explored.

A point of confusion rather than a potential design challenge is the change in
definitions between the IOD layer and the DAOS layer.  For the IOD layer, a
container is a collection of objects. For the DAOS layer, a container is a
collection of shards. For the IOD an object may be a shard of a global array.
For DAOS, a shard can host a set of DAOS objects. Having the same names with
locally correct, globally conflicting definitions serves to confuse how the
system should work.

\section{Conclusions}
\label{sec:conclusion}

The Fast Forward Storage and IO Stack project has designed a good first pass at
addressing the requirements for an extreme scale data storage mechanism. The
split between the IOD layer and the DAOS layer offers a fast place for
intermediate data without requiring the overhead of writing to persistent
storage. The envisioned transaction mechanism, while not perfect in the current
form, is another good attempt to address both failures and prevent access to
incomplete or incorrect data by downstream data consumers. Integrated with the
IOD functionality, this concept represents the consensus approach for what will
be required.

The partial metadata management incorporated into the IOD layer and the lack of
consideration for how to handle and recover from failures are oversights in the
current documents. It is our understanding that these will be addressed in the
next phase and we hope to help inform that effort with our experiences.

We hope that the efforts made in the \DDTns, Lightweight File Systems, and
other efforts to explore the requirements for this space, along with the
analysis presented in this paper will prove useful for the next phase of the
Fast Forward project.

\section{Acknowledgements}
\includegraphics[scale=0.07]{logos/doe_logo}
\includegraphics[scale=0.30]{logos/snl_logo}
\includegraphics[scale=0.35]{logos/nnsa_logo}
Sandia National Laboratories is a multi-program laboratory managed and operated
by Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
Corporation, for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-AC04-94AL85000.

\bibliographystyle{abbrv}
\bibliography{paper}

\vfill\eject

\end{document}