### 2014-04-28 paper.tex
IASDS submission version

\documentclass[conference]{IEEEtran}
\pdfpagewidth=8.5in
\pdfpageheight=11in
%\usepackage{subfig}
\usepackage{subfigure}
\usepackage[pdftex]{}
\usepackage{todonotes}
\usepackage{listings}
\lstset {% general command to set parameter(s)
         language=C,
     basicstyle=\footnotesize,               % print whole listing small
%     keywordstyle=\color{black}\bfseries, % underlined bold black keywords
%     identifierstyle =\color{black},  % nothing happens
%     commentstyle=\color{black}\emph, % white comments
     %stringstyle=\ttfamily,          % typewriter type for strings
%     stringstyle=\color{black},       % typewriter type for strings
         tabsize=4,
         showtabs=false,
     showstringspaces=false}
%can't figure this one out for particles bandwidth
%\usepackage[caption,label]{subfig}

%\newcommand{\TITLE}{D\textsuperscript{\huge{2}}T: Doubly Distributed Transactions for High Performance and Distributed Computing}

\newcommand{\DDT}{D\textsuperscript{2}T~}
\newcommand{\DDTns}{D\textsuperscript{2}T}

\hyphenation{sub-trans-ac-tion}

% change the inter-paragraph spacing
%\addtolength{\parskip}{-0.06in}

% in order for balance columns to work, it has to be before begin document...
%\balancecolumns

\begin{document}

%\conferenceinfo{MSST'14,} {November 17, 2013, Denver, Colorado, USA.}
%\CopyrightYear{2014}
%\crdata{978-1-4503-1103-8/11/11}
%\clubpenalty=10000
%\widowpenalty = 10000

\title{Consistency and Fault Tolerance Considerations for the Next Iteration of the DOE Fast Forward Storage and IO Project}

\author{
\IEEEauthorblockN{Jay Lofstead}
\IEEEauthorblockA{Sandia National Laboratories\\gflofst@sandia.gov}
\and
\IEEEauthorblockN{Ivo Jimenez}
\IEEEauthorblockA{University of California, Santa Cruz\\ivo@cs.ucsc.edu}
\and
\IEEEauthorblockN{Carlos Maltzahn}
\IEEEauthorblockA{University of California, Santa Cruz\\carlosm@soe.ucsc.edu}
}
\maketitle

\begin{abstract}
The DOE Extreme-Scale Technology Acceleration Fast Forward Storage and IO Stack
project is going to have significant impact on storage systems design within
and beyond the HPC community. With phase 1 of the project complete, it is an
excellent opportunity to evaluate many of the decisions made to feed into the
phase 2 effort. With this paper we not only provide a timely summary of
important aspects of the design specifications but also capture the underlying
reasoning that is not available elsewhere.

The initial effort to define a next generation storage system has made
admirable contributions in architecture and design.  Formalizing the general
idea of data staging into burst buffers for the storage system will help manage
the performance variability and offer additional data processing opportunities
outside the main compute and storage system. Adding a transactional mechanism
to manage faults and data visibility helps enable effective analytics without
having to work around the IO stack semantics. While these and other
contributions are valuable, similar efforts made elsewhere may offer attractive
alternatives or differing semantics that could yield a more feature rich
environment with little to no additional overhead. For example, the Doubly
Distributed Transactions (\DDTns) protocol offers an alternative approach for
incorporating transactional semantics into the data path. Another project,
PreDatA, examined how to get the best throughput for data operators and may
offer additional insights into further refinements of the Burst Buffer concept.

This paper examines some of the choices made by the Fast Forward team and
compares them with other options and offers observations and suggestions based
on these other efforts.  This will include some non-core contributions of other
projects, such as some of the demonstration metadata and data storage
components generated while implementing \DDTns, to make suggestions that may
help the next generation design for how the IO stack works as a whole.

\end{abstract}

%\category{D.4}{Software}{Operating Systems}
%\category{D.4.7}{Operating Systems}{Organization and Design}[hierarchical design]

%\terms{Design, Performance}

\section{Introduction}

Current production HPC IO stack design is unlikely to offer sufficient features
and performance to adequately serve the needs of an extreme scale platform. To
address these limitations, a joint effort between the US Department of Energy's
Office of Advanced Simulation and Computing and Advanced Scientific Computing
Research commissioned an effort to develop a design and prototype for an IO
stack suitable for the extreme scale environment. This is a joint effort led by
Lawrence Livermore National Laboratory, with the DOE Data Management Nexus
leads Rob Ross and Gary Grider as coordinators and contract lead Mark Gary. The
participating labs are LLNL, SNL, LANL, ORNL, PNL, LBNL, and ANL. Additional
industrial partners contracted include the Intel Lustre team, EMC, DDN, and the
HDF Group. This team has developed a specification
set~\cite{fastforward:2014:docs} for a future IO stack to address the
identified challenges. The first phase recently completed with a second phase
currently getting underway as of early 2014. The core focus of the first phase
was basic functionality and design. Overriding many of the decisions during
this and any subsequent phases is the reality of budgets. Placing GBs of NVRAM
on every node, while a potentially advantageous approach, is not financially
feasible. With this in mind, the second phase will refine this design
incorporating fault recovery and other features missing from the first phase.

The overall design seeks to offer high availability, byte-granular,
multi-version concurrency control. Through the use of a copy-on-write style
mechanism, multiple versions of an object can be stored in potentially greatly
reduced space. It assumes the client interface will be through an IO library
affording a more complicated interface that offers richer functionality
requiring only minimal end-user code changes.  Managing most data access in a
platform-local layer rather than requiring writing to centralized storage will
better support the performance and energy requirements of extreme scale
integrated application workflows. At a more detailed view, the various layers
of the IO stack each contribute different functionality and performance
implications.

The overall target machine architecture is illustrated in
Figure~\ref{fig:architecture}.

\begin{figure}[ht]
%\vspace{-0.15in}
\centering
\includegraphics[keepaspectratio=true, width=\columnwidth]{images/architecture}
%\vspace{-0.15in}
\caption{Target Machine Architecture}
\label{fig:architecture}
%\vspace{-0.15in}
\end{figure}

The basic architecture incorporates five layers. The top layer is a high level
IO library, such as the demonstration HDF-5 library~\cite{hdf5}. The intent is
to only have access to the storage stack through such an API to manage the
complexity of working with the lower layers. Below the user API is an IO
forwarding layer that redirects IO calls from the compute nodes to the IO
dispatching layer.  This IO forwarding layer is analogous to the function of
the IO nodes in a BlueGene machine. The next two layers have considerable
functionality. The IO dispatcher (IOD) serves as the primary storage interface
for the IO stack and is the only way to reach the persistent storage array in
lower layers. The Distributed Application Object Storage (DAOS) layer serves as
the persistent storage interface and is intended to be the foundation on which
everything else is built with no dependence on any technologies specified above
it. For example, the IOD layer and burst buffers are not required for DAOS to
operate properly. At the bottom is the Versioning Object Storage Device (VOSD).
It serves as the interface for storing objects of all types efficiently.

The focus of this paper is primarily the IOD layer given the critical role it
has in the performance and functionality of the entire stack. Most of the key
features explored in this paper all have a strong presence in the IOD layer
motivating the focus of this examination.

The core idea for IOD is to provide a way to manage the IO load that is
separate from the compute nodes and the storage array. Communication intensive
activities, such as data rearrangement, can be moved to the IOD layer
offloading the communication load from the compute nodes. IOD has three main
purposes. First, if the optional burst buffer is available, it works as a fast
cache absorbing write operations for the slower trickle out to the central
storage array. It can also be used to retrieve objects from the central storage
array for more efficient read operations and offers data filtering to make
client reads more efficient.  Second, it offers the transaction mechanism for
controlling data set visibility and to manage faults that would prevent a data
set from being used. Third, data processing operations can be placed in the
IOD. These operations are intended to offer data rearrangement, filtering, and
similar operations prior to data reaching the central storage array.

While these ideas are not necessarily new, they are new twists on best of class
efforts for these technologies. For example, offloading the collective
two-phase data sieving from the compute nodes to reorganize data has proven
effective at reducing the total time for writing data due to fewer participants
involved in the communication patterns~\cite{lofstead:2011:nessie-staging}.
Beyond these broad items, there are many important details some of which are
examined in more detail below.

Along with the analysis of the published design documents, a discussion of the
design philosophy representing the overall intent is presented. This
information represents information that may or may not have been written down,
but is the intent of ultimate product. These insights were gained based on
personal conversations with the team
members~\cite{JohnBent,QuinceyKoziol,EricBarton} discussing some of the
potential challenges with the design as written. These ideas are presented to
give a fuller picture of where the project is going rather than dwelling on the
limitations of the published documents.

The rest of the paper is organized as follows. First, an overview of the
target machine architecture and how it maps to the proposed components is
presented in Section~\ref{sec:overview}. Section~\ref{sec:burst} provides a
more detailed overview of the IOD layer and discusses some of the features of
incorporating burst buffers as designed and suggests some considerations and
alternatives for the next generation of this project.
Section~\ref{sec:transactions} discusses the transactions approach offered in
the IOD layer and the corresponding epochs in the DAOS layer. It also offers a
comparison to the \DDT system given the very similar high-level design and
motivating use case.  Section~\ref{sec:summary} discusses the system overall
with recommendations on what design elements should be considered based on
broader issues with current HPC data centers. The paper is concluded in
Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
paper.

\section{Overview}
\label{sec:overview}

The demonstration system uses some particular existing technology to enable
testing the proposed concepts. The mapping is presented in
Figure~\ref{fig:arch-mapping}. We examine this mapping by evaluating the
descriptive blocks at the bottom of the figure.

\begin{figure}[ht]
%\vspace{-0.15in}
\centering
\includegraphics[keepaspectratio=true, width=\columnwidth]{images/arch-mapping}
%\vspace{-0.15in}
\caption{Mapping FFSIO Components to Machine Architecture}
\label{fig:arch-mapping}
%\vspace{-0.15in}
\end{figure}

The dark blue block that includes the HDF5 VOL block represents the User API
layer. It is worth noting that while this is the intended interaction
mechanism, it is still possible for other interfaces like MPI-IO and POSIX to
interact with the system. The IO Forwarding layer is represented by the two
black boxes. The represented split shows that the IO nodes can offer some
control over how data is moved from the compute area towards storage. The
IO dispatcher, the primary focus of the evaluation presented in this paper, is
shown in the light green box. The one extension is the dark pink block
representing technology like SSDs on the IO nodes. This is where a burst
buffer, if the NVRAM is available, would be hosted. The DAOS layer is the
yellow box. The VOSD is the purple box.

The specification of HDF5 and Lustre at the top and bottom of the stack is 
an expedience for demonstrating the functionality. In a production version of
the proposed system, these may be represented by other technologies.

Throughout the paper, multiple terms are used to define an operation applied to
data outside of the main application computation. These include both function
shipping and just the term operator. In all cases, the idea is a piece of code
is placed at some location along the IO path processing data as it passes
through that point. Note that this works in both directions as well as in
place. For example, when reading, an operator may be used to only retrieve the
pieces of a large data set from the DAOS and VOSD layers into the IOD caches
automatically.  It is also possible that an operator is deployed to either
rearrange data or generate a new copy with different ordering. For example,
changing the fast dimension for a multi-dimensional array can greatly improve
reading performance for particular reading patterns. In all cases, these
operators offer a chance to apply a relatively light piece of processing to
data rather than an intensive data analytics operation. These heavy-weight
operations should be left to the main computation resources leaving capacity
for other concurrent applications using the IOD layer.

\section{IOD Layer and Burst Buffers}
\label{sec:burst}

The IOD layer is intended to function as a higher performance, machine local
caching layer for the DAOS/VOSD layers. Even in a simplified system, by keeping
the IOD layer as part of the stack, making portable code is simpler since it
eliminates the need to use a different lower level APIs from the user layer
(e.g., HDF5) if the system lacks a feature, such as burst buffers. In an ideal
design, the IOD API could function solely an directly on the DAOS servers. The
reality is that performance demands and budget constraints will force something
like burst buffers to compensate for the performance afforded and to offer a
location to process data down to volumes that can fit through the limited
bandwidth. Given the criticality of burst buffers and the transaction mechanism
for end-to-end performance, these ideas are examined in detail below with
comparison to alternatives that may have been avoided due to perceived
performance limitations. In particular, the \DDT protocol was designed to
address these sorts of scenarios and has demonstrated excellent performance.

The idea of burst buffers were initially explored in the context of data
staging~\cite{abbasi:2007:datatap,Abbasi:2009:datatap,nisar:2008:staging,zheng:2010:predata}.
These initial designs all use extra compute nodes to represent the data storage
buffer given the lack of any dedicated hardware support for this functionality.
The desired outcome of these initial studies is to motivate how such
functionality might be incorporated and the potential benefits.  Later, these
concepts were proposed to be incorporated as part of the IO
stack~\cite{nowoczynski:2008:zest,bent:2012:challenges,bent:2012:burst-buffer}.
The current Fast Forward IOD design recommends incorporating SSDs, but
specifically lists these devices as optional. Unfortunately, not incorporating
burst buffers and the use of SSDs in the IOD layer may be problematic.  First,
the IOD design currently is written assuming burst buffers. This means that the
bulk of IO operations will only hit the IOD layer and proposed functionality,
such as the function shipping, do not discuss the impact on the DAOS layer or
the operators themselves should a burst buffer not be available.  Consider the
important functionality of data rearrangement and doing things like changing
the fast array dimension on shared, spinning media.  The design of DAOS assumes
that it will only be involved when persisting a completed transaction and only
for a fraction of the total transactions created.  Transaction frequency at the
IOD layer can be higher since IOD does not ``flatten'' transactions. Flattening
is the process of transferring the set of copy-on-write changes from
transaction/epoch $n$ to transaction $n+m$ into the DAOS layer. The DAOS layer
mirrors the copy-on-write functionality, but stores the differences between
epochs rather than individual transactions. The primary difference is that at
the DAOS layer, data is stored in large, logically contiguous chunks to manage
the metadata load and hopefully improve read performance.

One of the bigger concerns is that the original data staging proposals all used
compute nodes while the newer proposals seek not only to make them a fixed
portion of the IO stack, but also shared across all machine users. The
PreDatA~\cite{zheng:2010:predata} paper in particular examines the potential
costs and advantages of where to place operators similar to the IOD proposed
function shipping. There are two key takeaways from PreDatA. First, placement
matters.  Depending on the communication intensity vs. computation intensity,
where along the IO path to place the operation can matter significantly.
Second, and more importantly, the amount of time spent processing for the
operators was stretched to the point where it consumed nearly all of the time
between IO operations. The given ratios of compute processes to staging process
examined is representative for future extreme scale platforms. If anything, the
ratios offer more staging processes than IOD processes would be available.

In the case of the written IOD design, it describes a fixed-sized staging area
that is partitioned on a per-application basis.  This is unlikely to be useful
because of the limited compute and communication capacity to spare to perform
these operations at a bottleneck in the IO path.  The use of a separate data
staging area intentionally separate from the IO path allows using operators on
limited resources leaving the IO path clear for strictly data movement. A
nuance of this design is discussed in the Design Philosophy below.

By concentrating the Burst Buffers and function shipping into the storage
stack, three problems arise.  First, the amount of network bandwidth, IO
bandwidth, and compute power consumed for example operations from a single
application is likely to completely monopolize the IOD processes. Second, if
space and time partitioning is used instead, the functionality risks being too
small to be useful. For example, if some of the IO nodes are dedicated either
throughout an application run (space partitioning) or for just phases of the
application runtime (time partitioning), the allocated capacity or ownership
duration may be insufficient to be useful. Third, the long-term hardware
performance advantage for SSDs is questionable. Recent studies have shown that
the erase-before-write and interference between reading and writing with
flash-based SSDs can cause severe performance
problems~\cite{skourtis:2013:ssd-performance}. The inclusion of an optional use
SSD layer in the new Trinity machine at Los Alamos will offer a test bed to
determine how likely these observed problems would affect a production extreme
scale platform.

Current NAND-based flash devices top out at around 400 MB/sec.  The key spec
that is missing from this number is that 400 MB/sec is a measure of the fixed
number of available IOPS multiplied by the block size. This represents the
ideal streaming performance possible. The problem is that it costs an IOP to
read 1 byte or 1 block (4KB or 8KB, depending on the device).  It costs 1 IOP
to write a full block--usually. In some cases, it will cost 2 IOPs.  This
accounts for the required pre-erase write prior to writing to a reused block.
In the worst case, it can be 3 or more IOPS per write. For example, if writing
forces both garbage collection and block compaction, many blocks may have to be
read, the data combined, the blocks erased, the combined data written, and then
the intended write operation can finally proceed on the newly freed blocks
(erase and then write).  One-third of 400 MB/sec, about 133 MB/sec, is well
below the streaming performance of HDDs.  Granted, there is still rotational
and seek latency to deal with for HDDs, but the advantage for SSDs has
evaporated and potentially turned into a considerable penalty at a cost
premium.  There are faster SSD solutions on the market that incorporate DRAM
for caching and using the PCIe bus, for example, but their price precludes them
from use in an extreme scale platform given budget constraints.

Given these features, the optionality and even incorporation of burst buffers
in the current design should be carefully considered. Much of the advanced, key
functionality proposed as they are currently designed ultimately relies on the
existence of burst buffers to work. Further thought about how to have an IOD
layer both with and without a burst buffer is required before they can be
considered optional. As the design stands today, they are a required part of
the IOD layer for proper functioning. Unfortunately, it is not clear that they
can address the performance concerns they are intended to cover.

\subsection{Design Philosophy}
The burst buffers design, as presented in the IOD documents, limits the
placement of the function operators and SSD buffers to the IO nodes. The team
does acknowledge the limitations of this design and intend to ultimately focus
on spreading the IOD layer from the IO nodes into the compute area as well.
This is intended both to help address the limitations of the IO bandwidth and
compute capability of these few nodes for data processing, but also to take
advantage of new layers in the storage hierarchy. By incorporating NVRAM into
compute nodes, new options for buffering data prior to being moved to
centralized storage become available and addresses some of the concerns about
SSD performance. For example, including a small amount of Phase Change memory
into many or most compute nodes offers a way to move data outside of both the
compute and IO path for data and communication intensive operations. Other 
projects~\cite{zheng:2010:predata} have suggested this will have value, but the
cost will have to be considered as part of the overall platform budget. This
lessens the impact of some operators while offering additional options for
places to store data.

Burst buffers being optional is a high level goal, but not one considered at a
detailed level within the design. For example, if there is no burst buffer, all
of the advanced functionality proposed for the IOD layer would have to work
against the DAOS layer instead. For example, function shipping assumes it will
operate on fast, local data within the IOD layer rather than against the
globally shared DAOS layer. With the additional desire to support using compute
node resources for these operations, serious work will be required to make a
fully functional end-to-end IOD layer implementation for a production system.

Another concern that is acknowledged, but no thought has been applied to, is
the requirement that a single IOD process of the set assigned to an application
is the master for any operation.  Should the number of concurrent applications
exceed the available nodes, sharing an IOD process will be required. The
requirements both in terms of scheduling and resource management were cut from
the project due to funding limitations. Since partitioning of the IOD processes
for exclusive use by particular applications is the assumed operating mode,
should insufficient IOD resources be available, either a job could be delayed
or IOD resources could be reallocated from a different process could be
redeployed for use by the new job. Handling resilience concerns for the IOD
processes must also be address.  These sorts of considerations still need to be
made for a full production system.

\section{Transactions, Epochs, and Metadata}
\label{sec:transactions}

The transaction mechanism manifests in two forms. At the IOD layer, they are
called transactions and are used to judge whether or not a set of distributed,
asynchronous modifications across a set of related objects is complete or not.
It is also used to control access by treating the transaction ID of committed
transaction as a version identifier.  At the DAOS layer, they are called epochs
and represent persisted (durable) transactions from the IOD layer. Each of
these offers different functionality, but are connected as is explained below.
How these differ from the \DDT approach is also explored.  While IOD's and
\DDTns's transactions are seemingly very different, they use a similar
high-level design, but very different implementation, to solve the same
problem.

\subsection{IOD Transactions}
To understand how transactions are used in the IOD layer, some terminology and
concepts must be explained first. At the coarsest grain level is a container.
Each container provides the single access context through which to access a
collection of objects. Transactions are used to treat a series of modifications
to the objects within a container atomically. Conceptually, containers
corresponds to a something akin to an HDF5 file in a traditional file system.
The objects in each container represent different data within a file.  The
three initially defined object types are key-value stores, multi-dimensional
arrays, and blobs.  The easiest way to understand these types is to evaluate
these from the perspective of an HDF-5 file, the initial user interface layer.
The key-value store represents a collection of attributes or groups. The array
represents a potentially multi-dimensional array.  The blob represents a byte
stream of arbitrary contents.  The fundamental difference between an array and
a blob is that the array has metadata specifying the dimension(s). At the
physical layer within the IO nodes, all of these objects may be striped across
multiple IO nodes.  Given this context, the transactions come in two forms.

First is a single leader transaction managed by IOD based on calls from a
single client. The underlying assumption is that the client side will manage
the transactional operations itself and the single client is capable of
reporting to the IOD how to evolve the transaction state. Passive failures are
detected based on the single client only. Any actively detected failure will
have to be managed by this single client.

The second form is called multi-leader and has the IOD layer manage the
transactions from a collection of clients. In this case, when the transaction
is created, a count of clients is provided to the IOD layer. As clients commit
their changes to the container, the reference count is reduced. Once the count
reaches 0, the transaction is automatically committed. Failure detection is
fully passive based on all participating clients.

\subsubsection{Design Philosophy}
Undocumented, but inherent in the design of these transactions is how faults
are detected. The initial design assumes the current Lustre fault detection
mechanism that can determine if a process or node is no longer reachable. This
detection happens at the DAOS layer and when a fault is detected, the rollback
process is pushed up to the IOD layer for all non-persisted or non-committed
transactions. This defines how a fault will be detected and what will trigger a
passive fault recovery (i.e., transaction abort).

There are two steps for beginning a transaction on a container. The first step
is for one or more process to open the container. This handle can be shared
eliminating the need for every participating process to hit the IOD layer to
open the file. The second step is a call to determine how many leaders will
participate in the transaction. In the single leader case, there is no
aggregation of success/fail statuses to determine the final transaction state.
Instead, it is assumed that the client will fully manage the transaction. In
the multi-leader model, some subset from 2 to $n$ where $n$ is the count of all
processes, declare themselves a leader for this container operation to the IOD
layer. Any number of processes can participate in modifying container without
regard to whether or not they are a leader. Once each leader has finished, with
the assumption that any clients they may be responsible for are finished as
well, the IOD layer aggregates those responses to either commit or abort the
transaction.

Ultimately, with the passive detection of faults for transaction leaders, the
transaction mechanism can work very well. A mostly unstated restriction that is
being relaxed is that every sequential transaction on a container is considered
dependent on the earlier transaction. Should one output be delayed and the
subsequent five succeed, when the delayed process finally fails, all six
transactions are rolled back. The thought of using this mechanism to store
subsequent checkpoint outputs in the same container to both save space, but not
care if one fails, cannot work in the current form. This has been acknowledged
and is being relaxed requiring a new parameter to the creation of a transaction
determining if it will be dependent or not.

\subsection{DAOS Epochs}
The Epoch mechanism differs from transactions. Instead of focusing on when a
particular output is complete, an epoch represents incremental persisted copies
of a container. To simplify the mapping between an IOD transaction and the DAOS
epochs, when an IOD transaction is persisted to DAOS, the IOD transaction ID is
the used as the epoch ID. The key difference is that at the DAOS layer, some
transaction (epoch) IDs will not be represented since not all IOD transactions
are necessarily persisted.

\subsection{Metadata Management}
Metadata management has been a perennial challenge for parallel storage
systems.  Eliminating metadata management as a special case and instead
treating it just as data is a central design goal of the Fast Forward project.
This is a hybrid approach to metadata management that is half-way between
providing no inherent metadata support and having a fully integrated, but
separate metadata management system.

Eliminating metadata as a core component of a file system is not new. It has
been explored as part of the Light Weight File Systems
project~\cite{oldfield:lwfs}. In LWFS, the metadata service is explicitly
limited to a user task with the storage layer limited to data
storage/retrieval, authorization, and authentication. This approach proved
workable. Using this hybrid approach is less common~\cite{weil:2006:ceph} and
introduces other issues.

IOD and DAOS both share a philosophy that they will have to maintain the
metadata about how the physical pieces of the logical objects are striped and
where they are placed. The primary metadata management is done at the DAOS
layer with the IOD layer relying on the DAOS layer for all authoritative
information about containers and objects. The only place where the IOD layer
manages metadata for itself is to manage how the different objects are striped
across the IO nodes.

\subsubsection{Design Philosophy}
While the metadata design is not fully defined, there are a few things that
are intended. For example, there is a standard, well-known container that is
the system metadata. This includes the list of all other containers. This
container is treated like any other data in the system and striped as
appropriate. Unfortunately, this still couples the metadata to a single object
that must serialize access. If the metadata, including information about
striping and other data layout operations were separated completely from the
data path, more scalable throughput could be achieved. The real challenge of
this is introduced by the IOD, DAOS, and VOSD layers collectively. Each of
these requires some different metadata storage and the migration is transparent
to the user.  Supporting fully independent metadata with this model is
difficult. Serious thought on how to do this effectively outside the data path
should be considered for phase two.

Based on the lessons from the \DDT metadata
service~\cite{lofstead:2012:txn-metadata} construction and the prior
experiments with LWFS, having a completely separate metadata service is
workable. Rather than making it a bottleneck in the IO path, it is another
service that users must interact with if they need those services.  Users can
manage everything by maintaining the metadata including the list of objects
themselves. However, there are drawbacks to this 100\% client-side approach.

With a client-side only approach, there is a serious risk of the metadata
service and the object store getting out of sync.  While having a metadata-less
object storage service is desirable, the different semantics from traditional
file systems requires some considerations. In this case, should these services
get out of sync, three particular risks are introduced.  First, a client could
create a dangling entry in the metadata service that does not correspond to any
objects in the object store. Second, a client could create orphaned objects
that have no associated metadata entries. Third, updates to the metadata or
object store service should be an atomic operation, but due to a lack of
coordination, a window where the system is inconsistent appears.

Ultimately, the consistency semantics required must be determined. If a
metadata service is required and it must be in sync with the object storage
service, then additional work must be performed. In traditional file systems,
the metadata and object storage updates are atomic. With decoupling metadata
from object storage, should this atomicity still be desired, it requires both
the ability for the services to participate in a task that is part of a larger
atomic operation and a higher-level mechanism to manage the atomic operation.

Overall, while additional work is required to maintain a client-side only
metadata service, it eliminates any potential bottlenecks related to updating
metadata related to the object storage. The burden of tracking striping and
other metadata that has traditionally been part of the metadata associated with
the file system will have to be maintained by the object storage service. The
lack of a centralized, serialized bottleneck to store that information improves
concurrency.

\subsection{Comparison to Other Protocols}
Alternatives, such as Paxos~\cite{Lamport:1998:paxos} algorithms like
ZooKeeper~\cite{Hunt:2010:zookeeper}, suffer from two limitations making them
unsuitable for this environment. First, the distributed servers in Paxos
systems are all distributed copies of each other that eventually become
consistent. Given the scale we wish to address, a single node's memory is
unlikely to be able to hold all of the data necessary for many operations at
scale. They also do not have a guarantee for when consensus will be achieved
without using slower synchronous calls. For the tight timing we wish to
support, we need guarantees of when a consistent state has been achieved.
Second, these systems also all assume that updates are initiated from a single
client process rather than a parallel set of processes as is the standard in
HPC environments.

Another effort to offer consistency and data integrity for the ZFS file
system~\cite{zhang:2010:zfs} covers some of the same territory. Instead of a
focus on the processes all having a notion of completion as a transaction, this
work focuses on the integrity of the data movement operations. We view this
work as something that should be considered hand-in-hand with a transaction
approach to ensure the integrity of the movement of the data in addition to the
agreement of processes about the successful completion of a parallel operation.

The transactions and epochs approach offered by FFSIO is most similar to
\DDT, but is a specialized implementation of a similar protocol. In this case,
\DDT is general and must use more general techniques. The FFSIO transactions
and epochs take advantage of simplifications based on the architecture and the
very specific use.

\DDT uses a second layer of coordination on the client side that greatly
increases the scalability by consolidating messages from clients into unique
sets prior to sending to the overall coordinator. A gossip
protocol~\cite{ganesh:2003:gossip-protocols} may appear sufficient for this
purpose, but the delay of eventual consistency is strictly avoided with this
protocol to ensure guarantees at particular states in the code. For example, if
a failure occurs, the global understanding of the role of all processes is
required in order for effective communication to occur for operations like
creating sub-transactions or voting. In this case, the protocol can offer
stronger statements about consistency than these protocols offer.  These
features offer a way to easily scale the transaction protocol given the
guarantees we wish to offer. The IOD approach does nothing to address these
concerns.

The details of \DDT is discussed next to offer a performance evaluation that
may be slower than FFSIO transactions and epochs for synchronous due to the
explicit messaging rather than relying on the hardware mechanisms. At scale,
the FFSIO performance should be similar to the \DDT approach.

\subsection{Comparison to \DDTns}
The \DDT project~\cite{lofstead:2012:txn} sought to develop an efficient
approach for handling ACID-style transactions in an environment with parallel
clients and multiple servers (doubly distributed). Rather than being aimed
solely at data movement operations, \DDT seeks to address the general problem
of managing any operation with multiple clients and servers.  Consider the
management of the analysis/visualization area, potentially similar to the IOD
concept. The transaction protocol is used to help manage resizing of the
resource allocation to the various analysis and visualization components.  For
the purposes of this discussion, \DDT could also be used to manage changing how
IOD processes and/or nodes are used without exposing these changes to the
client processes prematurely.  This has been described and analyzed
previously~\cite{dayal:2013:io-containers}.

The example metadata and data storage services created as part of the \DDT
project have no dependencies between transactions that prevent visibility
should an older version be incomplete. This additional, intentional requirement
by IOD offers different functionality than \DDTns's example services. In the
case of \DDTns, the functionality is less, but also avoids some of the concerns
outlined below.

The second iteration of the protocol~\cite{lofstead:2013:pdsw-txn} fixed
scalability issues and demonstrated a scalable client-side coordination model
with excellent performance. The performance measured for a complex transaction
with \DDT is illustrated in Figure~\ref{fig:performance}. This performance is
explored in detail in a previous paper~\cite{lofstead:2013:pdsw-txn}.
The
breakdown of the number of participants in each role is shown in
Table~\ref{tab:scaling}. For comparison, consider the Number of
Sub-Coordinators equivalent to IOD processes. The Processes Per Sub-Coordinator
represents the number of clients that use a particular IOD process. For these
tests, we maintained a balanced distribution and always used at least two
sub-coordinators to slow down the processing.

\begin{figure}[ht]
%\vspace{-0.15in}
\centering
\includegraphics[keepaspectratio=true, width=\columnwidth]{images/performance}
%\vspace{-0.15in}
\caption{Total Transaction Overhead}
\label{fig:performance}
%\vspace{-0.15in}
\end{figure}

\begin{table}[ht]
    %\vspace{-0.15in}
    \centering
    \caption[Scaling Configuration]{Performance Tests Scaling Configuration}
    \bigskip
    %\vspace{-0.15in}

    \begin{tabular}{|r|r|r|}
\hline
Processes & \vtop{\hbox{\strut Number of}\hbox{\strut Sub-Coordinators}} & \vtop{\hbox{\strut Processes Per} \hbox{\strut Sub-Coordinator}}\\
\hline
%8 & 2 & 4 \\
%16 & 2 & 8 \\
%32 & 2 & 16 \\
%64 & 2 & 32 \\
%128 & 2 & 64 \\
256 & 2 & 128 \\
512 & 2 & 256 \\
1024 & 4 & 256 \\
2048 & 8 & 256 \\
4096 & 16 & 256 \\
8192 & 32 & 256 \\
16384 & 64 & 256 \\
32768 & 128 & 256 \\
65536 & 256 & 256 \\
\hline
    \end{tabular}
    \label{tab:scaling}
\end{table}

At a high level, both \DDT and the IOD transactions have the same design. In
both cases, a hierarchical model is employed. In the case of \DDTns, it is a
purely client-side tree using semi-synchronous messaging. The messaging itself,
in the current implementation, uses asynchronous MPI messages. The synchronous
component comes from the timeout mechanism used to detect faults.  It forces a
level of coordination and synchronization for the protocol. For IOD, it is a
server-side tree and fully asynchronous relying on the existing Lustre fault
detection mechanism for failure detection. In both cases, there is a master in
charge of managing the transaction and a collection of workers that aggregate
into the master through second-level leaders. Beyond that, there are some
significant differences. Some of the different choices made by IOD raise some
possible concern.

The multi-leader model introduces the possibility of forcing a rollback of an
entire transaction when a partial retry might be sufficient for success. Since
the transactions are managed at a high level rather than the individual
tasks, a failure in a limited distributed task can cause the entire transaction
to fail. For example, consider 10 processes each have 5 tasks, but 3 of those
10 have an additional shared task to complete. If the task shared by the 3
processes fails on any of the three, the entire transaction would roll back
because it is a coarse-grained success/failure. If a concept like
sub-transactions at a task granularity were used, then it would be possible for
the one process that failed to report just that failure. Then the transaction
manager could reassign the resources for these 3 processes and try just that
operation again. If it now succeeds, then the overall transaction can be marked
successful only redoing the minimum amount of work required.

\begin{figure}[ht]
\centering
\includegraphics[keepaspectratio=true, width=0.8\columnwidth]{images/optimized-protocol}
%\vspace{-0.15in}
\caption{Optimized Protocol}
\label{fig:optimized-protocol}
%\vspace{-0.25in}
\end{figure}

\DDT has addressed theses issues in a couple of ways. First, the
sub-coordinators each have a list of processes from which they expect messages.
Should a message be missed, it is noticed and corrective action can be taken.
Second, \DDT has the concept of sub-transactions. The messaging requirements
are illustrated in Figure~\ref{fig:optimized-protocol}.  Sub-transactions
represent finer grained operations than the entire output, \DDT can manage
multiple writes per client by using a sub-transaction to represent the output
for any item to the file (container). Because of how the sub-transactions are
managed, the singleton sub-transactions, ones in which only a single process
participates, must be declared before the transaction begins so that its
existence can be broadcast as part of the begin transaction message. This
ensures there is global knowledge that the sub-transaction is expected.  That
way if the coordinator (transaction leader) fails, whichever process takes over
that role knows to expect a completion message for that sub-transaction or the
overall transaction cannot complete. Global sub-transactions can be defined at
any time since they are a global, synchronized operation broadcasting their
existence. While this additional layer does introduce messaging, the overhead
is quite small.

The advantages of eliminating these messages is not performance as demonstrated
by the performance of \DDTns. Instead, it offers a much less synchronous model
that matches with different programming models, such as Charm++ or other
task-based approaches. Since it can work for the bulk-synchronous model also,
it is a more broadly applicable approach. This assumes that the observed
potential issues can be addressed successfully.

\section{Broader Design}
\label{sec:summary}

At a broader level, there are some concerns that were partially clarified
through conversations with the team.  Consider a shared file system across an
HPC data center. The current design maintains the metadata in its own
container. Since copying data from the IOD layer to the DAOS layer requires an
explicit persist call, how and when synchronizing the metadata across the
layers and potentially across machines occurs is undefined. Delaying
synchronization until an explicit persist is called will reduce the update
frequency, but delays the data visibility on other platforms. Ideally, the
metadata object would need to be automatically persisted every time a container
transaction is persisted to the DAOS layer.

The implication of this is that every transaction persist is double operation
to account for the metadata persist. More importantly, the IOD-layer version of
the metadata container may contain readable transactions that have not been
persisted to the DAOS layer. How to handle this inconsistency between the two
layers still needs to be explored.

A point of confusion rather than a potential design challenge is the change in
definitions between the IOD layer and the DAOS layer.  For the IOD layer, a
container is a collection of objects. For the DAOS layer, a container is a
collection of objects across a set of shards. For the IOD an object may be a
shard of a global array.  For DAOS, a shard can host a set of DAOS objects.
Having the same names with locally correct, globally conflicting definitions
serves to confuse how the system should work.

\section{Conclusions}
\label{sec:conclusion}

The Fast Forward Storage and IO Stack project has designed a good first pass at
addressing the requirements for an extreme scale data storage mechanism. The
split between the IOD layer and the DAOS layer offers a fast place for
intermediate data without requiring the overhead of writing to persistent
storage. The envisioned transaction mechanism, while not perfect in the current
form, is another good attempt to address both failures and prevent access to
incomplete or incorrect data by downstream data consumers. Integrated with the
IOD functionality, this concept represents the consensus approach for what will
be required.

The partial metadata management incorporated into the IOD layer and the lack of
consideration for how to handle and recover from failures are oversights in the
current documents. It is our understanding that these will be addressed in the
next phase and we hope to help inform that effort with our experiences.

We hope that the efforts made in the \DDTns, Lightweight File Systems, and
other efforts to explore the requirements for this space, along with the
analysis presented in this paper will prove useful for the next phase of the
Fast Forward project.

\section{Acknowledgements}
The authors would like to thank Eric Barton, John Bent, Gary Grider, and
Quincey Koziol for their early review comments and discussions that clarified
the details of the design, the intent of the design, and the future plans.

\includegraphics[scale=0.07]{logos/doe_logo}
\includegraphics[scale=0.30]{logos/snl_logo}
\includegraphics[scale=0.35]{logos/nnsa_logo}
Sandia National Laboratories is a multi-program laboratory managed and operated
by Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
Corporation, for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-AC04-94AL85000.

We wish to thank Eric Barton for the some of the figures used to illustrate
the system architecture.

\bibliographystyle{abbrv}
\bibliography{paper}

\vfill\eject

\end{document}